---
title: "Climate Change Impact Modeling in Yazgulom River Basin: Preparation of Model Data"
author: "Tobias Siegfried, hydrosolutions GmbH"
format: html
editor: source
date: "2025-05-2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This R Quarto document details the process of preparing model data for climate change impact modeling in the Yazgulom River basin. The script is designed to process climate data, perform necessary transformations, and generate inputs for hydrological modeling. This work is crucial for understanding potential future changes in water resources within the basin under various climate scenarios.

We only have monthly data available for the Yazgulom River basin. The script is structured to handle the specific requirements of this dataset:
- hydro model calibration as in the Suek River case study
- climate impact study as in the Vaksh River basin

# Change Log

This section documents the evolution of the script, highlighting key updates and improvements:

- **2024-02-08**: Initial version created, adapting the original code from the Taloqan River basin study. The script reverts to using CHELSA climate data instead of NSIDC data.
- **2024-02-10**: Continued development to produce a workable script for Adrian.
- **2024-02-11**: Implemented new download option for CHELSA V21 data from WSL servers, significantly improving data acquisition. Major script reorganization and validation of sections up to Section 4. Sections 5 (Bias Correction) and 6 await further refinement.
- **2024-02-12**: Fixed a bug in the discharge data concatenation process.
- **2024-02-17**: Established as the working version to serve as a base for the handbook.
- **2024-05-13**: Added glacier-related components to the analysis.
- **2024-06-20**: 
  - Resolved IPSL error and improved overall script consistency.
  - Enhanced BCSD (Bias Correction Spatial Disaggregation) methodology, now correctly applied to IPSL data.
  - Implemented numerous improvements and beautification of code.
  - Added analysis of modeling results (available in the corresponding script `zrb_climate_impact_analysis.qmd` within the R project `model_zarafshan_tobi`).
- **2024-06-21**: Identified significant issues with the previously calibrated model.
- **2024-06-23**: Strategized on model improvement:
  - Considered recalibration or switching to Bea's best 10-layer model.
  - Noted challenges with changing models mid-teaching (Adrian and Aidar using 9-layer model).
- **2024-06-24**: Began test runs using Beatrice's model for climate change impact modeling predictions, due to unsatisfactory results with the 9-layer model.
- **2024-06-25**:
  - Reran entire database creation to address coordinate inconsistencies, now consistently using `hru_utm`.
  - Integrated decadal discharge data into the database.
  - Added a cross-check quality control section.
- **2024-06-26**: 
  - Major update: Replaced future IPSL-CM6A-LR projections due to incorrect precipitation fields.
  - Updated precipitation fields in corresponding FUT-SIM folder and Documents/ca_large_files/.
- **2024-06-27**: Introduced option for precipitation rescaling of historical observation data.
- **2024-07-07 to 2024-07-14**: Continued refinement of precipitation forcing rescaling methodology.
- **2024-07-17**: Added glacier processing section, completing all processing sections for a comprehensive workflow script.
- **2024-08-08**: 
  - Computed BCSD files with rescaled precipitation for historical observations.
  - Ran hydrological model and produced output results for analysis.
  - Enhanced `plot_forcing_data` function to display figures in a four-panel layout.
- **2024-08-09**: Reorganized file output writing for better organization of climate forcing data.
- **2024-08-12**: 
  - Added expert FOF PNG file for integrating figures in Windows Word.
  - Planned to streamline code by sourcing functions to reduce duplication.
  - Improved climate scenarios palette for enhanced figure visualization.
- **2024-08-13**: 
  - Enhanced glacier processing methodology.
  - Recomputed RS Minerve simulations to include glacier melt in future projections.
- **2024-09-30**: Adapted script for Vaksh basin, incorporating improvements from the intermediate Suek River basin study.
- **2024-10-04**: 
  - CHELSA V21 pr data only available until and including 2019 currently for download. Dirk Karger will inform, when the new data is available post 2000.
  - Lapse rates are currently not available for Vaksh to validate CHELSA V21 preciptiation data. Let us see if Andrey can get this data.
  - Various smaller improvements and code streamlining.
- **2024-10-05**: 
  - Further code improvements and streamlining. The goal was to make it more general that we can apply our approach to any basin without having to change variable names and directory paths. Looks like we are good now.
  - The future philosophy is to only refer to the basin name in the parent folder and to create and store the R project at this level (and not in the code folder). This makes the lower level folder structure more general and easier to maintain. It also eases the pain of replication to the next river. The resulting directory structure is shown below...
  - Added section 1.5 for precipitation data import and visualization for them in section 5.4: Precipitation data comparison

XYZ_river_modeling_story/
│
├── XYZ_river_modeling_study.R.proj
├── 01_code
├── 02_data
├── 03_materials
├── 04_report-see_oneDrive
└── 05_models
    └─ forcing
    
- **2024-11-18**:
  - Code review by @Tobi. Changed glacier preparation data to ensure that our hist_obs period is now correctly from 2000 through 2019. The fut_sim period can then start in 2020. This is a deviation from usual per period assessments but is necessary due to the data availability in the case of Vaksh.
  - Introduce factors for hru plotting just to make the legend nice looking.
- **2024-11-19**:
  - Correcting fut_sim data preparation issues.
- **2024-11-20**
  - Setup and export basin info, including the basin area, basin code, basin name, and hypsometric curve data. This will be read into the hydro_model_airGR. qmd script.
  - More preparation towards full integration of airGR workflow in the climate assessment/impact chain.
  - Need to check Hypso data since I believe something is wrong.
- **2024-11-21**:
  - Code review by @Aidar. Changed "Hydrological Modeling" chapter location from before "GIS Data" to after "GIS Data", because "Hydrological Modeling" uses some directories that were defined in "GIS Data:.
  - Store glacier melt data for the calibration and validation period for importing and processing in the hydro modeling script.
  - Need to validate the hypsometric curve implementation since we get very small elevation values in the curve. What could be wrong?
- **2024-11-25**:
  - Add elevation band data layer height to Basin_Info$ZLayers list so that we can retrieve this information in the hydro_model_airGR.qmd script.
- **2024-11-28**:
  - Further streamlining the code and harmoniziation between the scripts. 
  - Now sourcing the functions to provide them as context to LLMs.
- **2024-11-30**:
  - Greatly improved and polished plotting of figures.
- **2024-12-03**:
  - Figures of Lapse Rates need to be improved.
- **2024-12-18**:
  - Glacier volume calculation using Erasov method added
- **2025-02-18**:
  - Preparing for sharing the code with the team.
- **2025-02-27**:
  - Small edits and changes.
- **2025-03-02**:
  - Small edits to reflect changes in the .ipynb file.
  - Also installed Quarto in the conda `myrrun` environment for further testing qmd-files with VSCode.
- **2025-03-03**:
  - Added a parameter configuration file to the code folder and the corresponding configuration loading.
- **2025-03-04**:
  - Corrected newly introduced error in Section 3 GLACIERS DATA
  - Made numerous beautification changes. It's slowly getting presentable.
- **2025-03-05**:
  - Another round of essential code streamlining and optimization.
# ==========
- **2025-03-05**:
  - Adaptation of script to Atbashy catchment
- **2025-04-25**:
  - Further tweaking of script and making sure that model calibration with monthly data works.
- **2025-04-27**:
  - Further improvements and polishing of the script.
- **2025-04-28**:
  - Many changes and adaptations to run a future ERA5 scenario only.
  - Script is working well now and is ready for the next steps.
# ==========
- **2025-05-01**:
  - Spin off to Yazgulom River basin. Here, we use the Atbashy 2025 modeling approach for calibration and the Vaksh case for the impact study.

# CONFIGURATIONS

This section loads a centralized configuration file and sets up key variables for the analysis. The configuration includes parameters for river/basin identification, file paths, time periods, climate models and scenarios, GIS parameters, elevation band parameters, visualization palettes, and processing options. This structured approach makes the code more maintainable and adaptable to different river basins.

```{r}
# Load the centralized configuration
library(pacman)
p_load(here, tidyverse, lubridate)

path = here::here("01_code", "config_param.R")
source(path)

# Access configuration parameters through the config structure:
# - config$project: River and basin identification
# - config$paths: Data directory locations
# - config$periods: Time periods for analysis
# - config$climate: Climate models and scenarios
# - config$gis: GIS parameters and file paths
# - config$hru: Elevation band parameters
# - config$colors: Color palettes for visualization
# - config$options: Processing options and thresholds

# For backward compatibility with existing code, define key variables
# from the configuration
river_name <- config$project$river_name
basin_name <- config$project$basin_name
gauge_code <- config$project$gauge_code
gauge_name <- config$project$gauge_name

# Path configuration
data_path <- config$paths$data_path
gis_path <- config$paths$gis_path
hist_obs_dir <- config$paths$hist_obs_dir
hist_sim_dir <- config$paths$hist_sim_dir
fut_sim_dir <- config$paths$fut_sim_dir
model_dir <- config$paths$model_dir
glacier_path <- config$paths$glacier_path
fig_rep <- config$paths$figures_path

# Time periods
hist_obs_start <- config$periods$hist_obs_start
hist_obs_end <- config$periods$hist_obs_end
hist_sim_start <- config$periods$hist_sim_start
hist_sim_end <- config$periods$hist_sim_end
fut_sim_start <- config$periods$fut_sim_start
fut_sim_end <- config$periods$fut_sim_end
calib_start <- as.Date(config$periods$calib_start)
calib_end <- as.Date(config$periods$calib_end)
valid_start <- as.Date(config$periods$valid_start)
valid_end <- as.Date(config$periods$valid_end)

# Climate models and scenarios
gcm_Models <- config$climate$models
hist_sim_models <- config$climate$hist_sim_models
fut_sim_models <- config$climate$fut_sim_models
gcm_Scenarios <- config$climate$scenarios
gcm_Models_Scenarios <- config$climate$model_scenarios
hydrology_run_scenarios <- config$climate$hydrology_run_scenarios
hydrology_run_Models_Scenarios <- config$climate$hydrology_run_model_scenarios
obs_freq <- config$climate$obs_freq

# Generated date sequences
hist_obs_dates <- config$dates$hist_obs_dates
hist_sim_dates <- config$dates$hist_sim_dates
fut_sim_dates <- config$dates$fut_sim_dates

# GIS parameters
utm42n <- config$gis$utm42n
crs_project <- config$gis$crs_project
basin_shape_file <- config$gis$basin_shape_file
gauge_shape_file <- config$gis$gauge_shape_file

# Elevation band parameters
dh_elband <- config$hru$dh_elband

# Visualization palettes
pastel_palette <- config$colors$scenario_palette
elevation_colors <- config$colors$elevation_colors
atbashy_colors <- config$colors$atbashy_colors

# Processing options
rescale_pr <- config$options$rescale_pr
streams_threshold <- config$options$streams_threshold
hydro_model_time_steps <- config$options$hydro_model_time_steps
sec_month <- config$options$sec_month
```

# FUNCTIONS

The document sources numerous functions from external files, making them available in the current environment. These functions handle various aspects of the workflow including plotting, data processing, hydrological calculations, and specialized operations for Central Asian rivers.

```{r}
# Load the here package
p_load(here)

# Standard packages
source(here::here("01_code", "functions", "plot_summary_stats.R"))
source(here::here("01_code", "functions", "plot_dev_norm.R"))
source(here::here("01_code", "functions", "extract_year.R"))
source(here::here("01_code", "functions", "post_process_hru.R"))
source(here::here("01_code", "functions", "process_rsm_data.R"))
source(here::here("01_code", "functions", "plot_forcing_data.R"))
source(here::here("01_code", "functions", "process_glacier_runoff.R"))
source(here::here("01_code", "functions", "prepare_glacier_data_mean_annual_runoff.R"))
source(here::here("01_code", "functions", "calculate_hypsometric_curve.R"))
source(here::here("01_code", "functions", "calculate_total_glacier_volume.R"))
source(here::here("01_code", "functions", "create_enhanced_subseries_plot.R"))
source(here::here("01_code", "functions", "plot_seasonal_diagnostics.R"))
source(here::here("01_code", "functions", "plot_discharge_timeseries.R"))

# Below are better versions of the relevant riversCentralAsia package functions
source(here::here("01_code", "functions", "generateSeqDates.R"))
source(here::here("01_code", "functions", "decadeMaker.R"))
source(here::here("01_code", "functions", "monDateSeq.R"))
source(here::here("01_code", "functions", "loadTabularData.R"))
source(here::here("01_code", "functions", "glacierVolume_Erasov.R"))
source(here::here("01_code", "functions", "posixct2rsminerveChar.R"))
source(here::here("01_code", "functions", "gen_HRU_Climate_CSV_RSMinerve_local.R"))
```

# 1 DATA

This major section handles data loading and processing across multiple domains:

- Discharge data from gauge 16076, Atschikomandi in Atbashy River
- GIS data including basin shapefiles, digital elevation models, and hypsometric curves
- In-situ precipitation data from various stations
- Glacier outlines from the Randolph Glacier Inventory v6.0
- Data export for hydrological modeling parameters
- Basin overview visualization with detailed topographical representation

## 1.1 Discharge

We work with data from the Gauge 16076 Atbashy Atschikomandi We have only monthly datasets.

```{r station_details}
#| message: false
#| warning: false

# Gauges
# Currently, we only work with the 17084 Gauge (Darband, Vaksh)

# Monthly data (Tabular data)
q_mon <- loadTabularData(fPath = here::here("02_data", "Discharge"), 
                         fName = paste0(gauge_code,"_Q.csv"), 
                         code = gauge_code, 
                         stationName = gauge_name, 
                         rName = river_name, 
                         rBasin = basin_name, 
                         unit = "m3/s", 
                         dataType = "Q")

q_mon |> head()

q_mon |> tail()
```

## 1.2 GIS Data

### Raster and Shape Files
```{r}
#| message: false
#| warning: false

# Packages
p_load(sf, terra, elevatr)

# GIS 
#Important naming convention. We assume that GIS-files adhere to the following naming convention:
#- Basin Shapefile: paste0(gauge_code,"_Basin.shp")
#- River Shapefile: paste0(gauge_code,"_River.shp")
#- Junctions Shapefile: paste0(gauge_code,"_Junctions.shp")
#- HRU Shapefile: paste0(gauge_code,"_HRU.shp")
# -DEM Raster: paste0(gauge_code,"_DEM.tif")

# Load shape file of river basin
basin_shp <- sf::st_read(file.path(gis_path, basin_shape_file), quiet = TRUE)
basin_utm <- sf::st_transform(basin_shp, utm42n)
basin_latlon <- sf::st_transform(basin_shp, crs_project)

# compute centroid of the basin
basin_centroid <- sf::st_centroid(basin_latlon)
# get the coordinates of the centroid
basin_centroid_coord <- sf::st_coordinates(basin_centroid)
# convert latitude to rad
lat_rad <- basin_centroid_coord[2] * pi / 180

# if the dem does not exist, download it
if (!file.exists(file.path(gis_path, "dem.tif"))) {
  dem_latlon <- elevatr::get_elev_raster(basin_shp, z = 9)
  # convert dem_latlon to spatRaster
  dem_latlon <- terra::rast(dem_latlon)
  # convert to UTM42N
  dem_utm <- terra::project(dem_latlon, utm42n)
  terra::writeRaster(dem_utm, file.path(gis_path, "dem.tif"))
  
} else {
  dem_utm <- terra::rast(file.path(gis_path, "dem.tif"))
}

# gauge location and elevation
gauge_shp <- 
  sf::st_read(file.path(gis_path, gauge_shape_file), quiet = TRUE)

gauge_coord <- gauge_shp %>% sf::st_coordinates()
gauge_Z <- terra::extract(dem_utm, gauge_shp)[2]
gauge_Z <- as.numeric(gauge_Z)
print(paste("Gauge elevation:", gauge_Z, "masl"))

# Calculate the basin area
basin_area_m2 <- sf::st_area(basin_latlon) |> units::set_units(m2)
basin_area_km2 <- basin_area_m2 |> units::set_units(km2)
print(paste("Basin area:", round(basin_area_km2), "km2"))
```

### Hypsometric Curve

```{r}
#| message: false
#| warning: false

# first mask the dem with the basin
dem_utm <- dem_utm %>% terra::mask(terra::vect(basin_utm))
dem <- terra::project(dem_utm, "EPSG:4326")
# save dem on disk
terra::writeRaster(dem, file.path(gis_path, "dem_atbashy_basin.tif"), 
                   overwrite = TRUE)


# Calculate hypsometric curve
hypsometric_results <- calculate_hypsometric_curve(dem_utm, relative_height = FALSE, plot = FALSE)

# Access the hypsometric curve results
curve_data <- hypsometric_results$curve_data  # DataFrame with all calculations

# save figure
ggsave(file.path(fig_rep, paste0(river_name,"_hypsometric_curve.pdf")), 
       plot = hypsometric_results$plot, width = 12, height = 8, units = "cm", dpi = 600)

# Plot to screen
hypsometric_results$plot
```


## 1.4 Glacier outlines (RGI v6)

Currently only working with RGI 6.0 (not 7.0). This could be updated at one point in time. However, Rounce et al. also work with v6, so there is no immediate need to do so.

```{r}
#| warning: false
#| message: false

# Randolph Glacier Inventory (RGI) version 6.0 for glacier outlines in the project region. 
rgi <- sf::st_read(file.path(glacier_path,"rgi_CA_domain_UTM42N.shp"), quiet = TRUE) |> 
    distinct(RGIId, .keep_all = TRUE)
rgi <- rgi |> mutate(glacier_area_m2 = as.numeric(st_area(rgi)))
glaciers <- rgi$RGIId  # List of glaciers

# Basin outlines
basins <- basin_utm

# add the column DnstNode from code to the basins
basins$DnstrNode <- gauge_code
basins <- basins |> mutate(basin_area_m2 = as.numeric(st_area(basins)))

# Intersect the basins with the glacier outlines to attribute each glacier to a basin. 
rgi_in_basins <- sf::st_intersection(rgi, basins)

# Basin boundaries may not be consistent with glacier outlines, i.e. some glaciers may end up split into two basins. To avoid for double accounting of the glacier runoff, we determine an area weight which will then be multiplied with the glacier discharge from PyGEM.
rgi_area_basins <- rgi_in_basins |> 
  mutate(area_intersected_m2 = as.numeric(st_area(rgi_in_basins)), 
         weight = area_intersected_m2 / glacier_area_m2) 
```

## 1.5 Exporting Data

Export Hydrological Modeling Parameters and Basin_Info Parameters

```{r}
Calibration_Validation_Period <- list(
  calib_start = calib_start,
  calib_end = calib_end,
  valid_start = valid_start,
  valid_end = valid_end,
  sec_month = sec_month,
  hydro_model_time_steps = hydro_model_time_steps
)

# store to disc
saveRDS(Calibration_Validation_Period, file.path(data_path, "Calibration_Validation_Period.rds"))

Basin_Info <- list(
  BasinCode = gauge_code,
  BasinName = river_name,
  BasinArea_m2 = basin_area_m2,
  BasinLat_rad = lat_rad,
  HypsoData = curve_data$elevation)

saveRDS(Basin_Info, file.path(data_path, "Basin_Info.rds"))
```

## 1.6 Basin Overview Plot

This is the code to produce a very polished publication figure.

```{r}
# load required packages
p_load(ggspatial)
p_load(ggnewscale)

# SINGLE DIRECTION HILLSHADE
# dem dataframe
dem_df <- as.data.frame(dem, xy = TRUE)
names(dem_df)[3] <- "alt"
# estimate the slope
sl <- terrain(dem, "slope", unit = "radians")

# estimate the aspect or orientation
asp <- terrain(dem, "aspect", unit = "radians")

# calculate the hillshade effect with 45º of elevation
hill_single <- shade(sl, asp, 
      angle = 45, 
      direction = 300,
      normalize = TRUE)

# convert the hillshade to xyz
hill_single_df <- as.data.frame(hill_single, xy = TRUE)

# MULTI DIRECTION HILLSHADE
# pass multiple directions to shade()
hillmulti <- map(c(270, 15, 60, 330), function(dir){ 
                    shade(sl, asp, 
                          angle = 45, 
                          direction = dir,
                          normalize = TRUE)}
  )

# create a multidimensional raster and reduce it by summing up
hillmulti <- rast(hillmulti) %>% sum()

# multidirectional
#plot(hillmulti, col = grey(1:100/100))

# convert the hillshade to xyz
hill_multi_df <- as.data.frame(hillmulti, xy = TRUE) |> 
  #rename last column to hillshade
  rename(hillshade = sum)

# Create custom breaks for elevation
elev_breaks <- seq(floor(min(dem_df$alt)/500)*500, 
                  ceiling(max(dem_df$alt)/500)*500, 
                  by = 500)

# Create the enhanced hillshade plot
pl_topo_glaciers <- ggplot() +
  # Add hillshade base layer
  geom_raster(
    #data = hill_single_df,
    data = hill_single_df,
    aes(x, y, fill = hillshade),
    show.legend = FALSE
  ) +
  scale_fill_gradientn(
    colors = grey.colors(100, start = 0.3, end = 0.9),
    guide = "none"
  ) +
  # Add new scale for elevation
  new_scale_fill() +
  # Add elevation layer with transparency
  geom_raster(
    data = dem_df,
    aes(x, y, fill = alt),
    alpha = 0.6
  ) +
  # Custom elevation color scale
  scale_fill_gradientn(
    colors = elevation_colors,
    breaks = elev_breaks,
    #labels = scales::label_number(suffix = " m"),
    # Only show labels for breaks divisible by 1000
   labels = function(x) ifelse(x %% 1000 == 0, paste0(x, " m"), ""),
    guide = guide_colorbar(
      direction = "horizontal",
      title.position = "top",
      barwidth = 15,
      barheight = 1,
      title = "Elevation (masl)"
    )
  ) +
  # Add basin boundary
  geom_sf(
    data = basin_latlon,
    fill = NA,
    color = "darkgrey",
    linewidth = 0.1
  ) +
  # Labels and title
  labs(
    title = paste(river_name, "River Basin"),
    subtitle = "Terrain Elevation with Hillshade"#,
    #caption = "Data source: ALOS WORLD 3D - 30m (AW3D30)\nHillshade: 45° elevation, 300° azimuth"
  ) +
  # Coordinate system and extent
  coord_sf(expand = FALSE) +
  # Enhanced theme
  theme_void() +
  theme(
    # Text elements
    text = element_text(family = "Arial", color = "grey20"),
    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 5)),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    plot.caption = element_text(size = 8, color = "grey50", margin = margin(t = 10)),
    
    # Grid lines
    panel.grid = element_blank(),
    
    # Legend formatting
    legend.position = "bottom",
    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 8),
    legend.margin = margin(t = 10),
    
    # Plot margin
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  ) +
  # Add scale bar and north arrow
  ggspatial::annotation_scale(
    location = "br",
    width_hint = 0.2,
    style = "ticks",
    text_family = "Arial"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tl",
    which_north = "true",
    pad_x = unit(0.2, "in"),
    pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_minimal()
  ) +
# Add glaciers shape
geom_sf(
    data = rgi_in_basins,
    fill = "#A5F2F3",
    color = "steelblue",
    alpha = 0.3  # Added some transparency
)
```

Now, we plot basin, glaciers and rivers.

```{r}
#| warning: false
#| message: false
#| echo: false

# load packages
p_load(whitebox)

# Set River threshold
streams_threshold <- 100 # 100 is a good value for a beautiful river network visualization

# First need to fill depressions in DEM
wbt_fill_depressions(
  dem = file.path(gis_path, "dem_basin.tif"),
  output = file.path(gis_path, "dem_basin_filled.tif")
)

# Create D8 flow pointer grid
wbt_d8_pointer(
    dem = file.path(gis_path, "dem_basin_filled.tif"),
    output = file.path(gis_path, "dem_basin_d8_pointer.tif")
)

# Calculate flow accumulation 
wbt_d8_flow_accumulation(
  input = file.path(gis_path, "dem_basin_filled.tif"),
  output = file.path(gis_path, "dem_basin_flow_acc.tif")
)

# Extract streams using a threshold
wbt_extract_streams(
  flow_acc = file.path(gis_path, "dem_basin_flow_acc.tif"),
  output = file.path(gis_path, "dem_basin_streams.tif"), 
  threshold = streams_threshold  # Adjust this threshold value as needed
)

# Convert raster streams to vector
wbt_raster_streams_to_vector(
  streams = file.path(gis_path, "dem_basin_streams.tif"),
  d8_pntr = file.path(gis_path, "dem_basin_d8_pointer.tif"),
  output = file.path(gis_path, "streams_basin.shp")
)

# Read streams
dem_crs <- st_crs(basin_latlon)
streams <- sf::st_read(file.path(gis_path, "streams_basin.shp")) |> 
  sf::st_set_crs(dem_crs)

# To vary the stream thickness based on downstream distance or flow accumulation, we need stream ordering. We can add this using WhiteboxTools' stream order functions. The final plot then looks like this.

# After creating the initial streams, calculate Strahler stream order
wbt_strahler_stream_order(
    streams = file.path(gis_path, "dem_basin_streams.tif"),
    d8_pntr = file.path(gis_path, "dem_basin_d8_pointer.tif"),
    output = file.path(gis_path, "dem_basin_streams_ordered.tif")
)

# Convert ordered streams to vector
wbt_raster_streams_to_vector(
    streams = file.path(gis_path, "dem_basin_streams_ordered.tif"),
    d8_pntr = file.path(gis_path, "dem_basin_d8_pointer.tif"),
    output = file.path(gis_path, "streams_basin_ordered.shp")
)

# Read and set CRS
streams_ordered <- sf::st_read(file.path(gis_path, "streams_basin_ordered.shp")) %>%
    sf::st_set_crs(dem_crs)

# Now plot with varying line width based on stream order
pl_topo_glaciers_streams <- pl_topo_glaciers +
geom_sf(
    data = streams_ordered,
    aes(linewidth = STRM_VAL),  # ORDER is the column name WhiteboxTools creates
    color = "blue",
    alpha = 0.3
) +
scale_linewidth_continuous(
    range = c(0.2, 0.5),  # Adjust min/max line widths
    guide = "none"  # Remove legend if you don't want to show it
)

# save figure
ggsave(file.path(fig_rep,"basin_topo_glaciers_streams.png"), 
  pl_topo_glaciers_streams, width = 10, height = 5, dpi = 600)

#plot figure
pl_topo_glaciers_streams <- pl_topo_glaciers_streams + 
  theme(text = element_text(family = "sans"))

pl_topo_glaciers_streams
```

# 2 ANALYSIS OF DISCHARGE

 Comprehensive analysis of discharge data at both monthly and daily time scales. This includes creating consistent time series by filling gaps, visualizing long-term trends, identifying seasonal patterns, analyzing cold/warm season discharge variations, and establishing calibration and validation periods for hydrological modeling.

## 2.1 Monthly Data

### Plot Time Series of Monthly Discharge

Plotting the monthly time series reveals the long-term trends and seasonal patterns in the discharge data. There are some funny patters in the 1970ies. We will only use the data from 1979 onwards from which CHELSA V21 data is available.

```{r}
#| warning: false
#| message: false

p_load(timetk)

q_mon <- q_mon |> mutate(value = data)

# Calculate long-term average by month
monthly_averages <- q_mon %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarise(avg_value = mean(value, na.rm = TRUE))

# Fill NA values with long-term average for that month
q_mon <- q_mon %>%
  mutate(month = month(date)) %>%
  left_join(monthly_averages, by = "month") %>%
  mutate(value = coalesce(value, avg_value)) %>%
  dplyr::select(-month, -avg_value)

# Create the plot with a publication figure of the discharge time series
discharge_plot <- plot_discharge_timeseries(
  data = q_mon,
  date_col = "date",
  value_col = "value",
  gauge_name = gauge_name,
  gauge_code = gauge_code
)

# save plot
ggsave(file.path(fig_rep, "discharge_timeseries.png"), 
       plot = discharge_plot,
       width = 10, height = 5)

# Display the plot
discharge_plot
```

### Seasonal Diagnostics

The seasonal diagnostics plot shows the mean, median, and interquartile range of discharge values for each month. This plot helps identify any seasonal patterns or trends in the data.

```{r}
#| warning: false
#| message: false
#| echo: false

p_load(patchwork)

# Example usage:
# Assuming your dataframe is called 'atbashy_data' with columns 'date' and 'value'
pl <- plot_seasonal_diagnostics(q_mon)

ggsave(file.path(fig_rep,"seasonal_diagnostics.pdf"), plot = pl,
       width = 10, height = 5)

# plot on screen
pl
```


### Calibration and Validation Periods

Calibration period: 1979 - 1990
Validation period: 1991 - 1996
Simulation Period: 2000 - 2023 (Note: This requires us to resort to ERA5 forcing data fetched via the data gateway).

```{r}
#| warning: false
#| message: false
#| echo: false

# params
plot_point_size <- 1

# Create the 'Period' column
q_plot <- q_mon

q_plot$Period <- ifelse(
  q_plot$date >= calib_start & q_plot$date <= calib_end, "Calibration",
  ifelse(
    q_plot$date >= valid_start & q_plot$date <= valid_end, "Validation",
    NA  # Assign NA to dates outside the defined periods
  )
)      

q_hist_plot <- ggplot(data = q_plot, aes(x = date, y = value, group = 1)) +
  geom_line(aes(color = Period), size = 0.5) +
  geom_point(data = subset(q_plot, Period == "Calibration"), 
             color = "black", size = plot_point_size) +
    geom_point(data = subset(q_plot, Period == "Validation"), 
               color = "red", size = plot_point_size) +
  labs(#title = "Calibration and Validation Periods",
       x = "Date",
       y = "Discharge [m3/s]",
       color = "Period") +
  theme_bw()  +
  scale_color_manual(values = c("Calibration" = "black", 
                                "Validation" = "red")) +
  theme(
    #plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
    #panel.grid.major = element_line(color = "grey90"),
    #panel.grid.minor = element_line(color = "grey90")
  ) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")

# save to figures repository
ggsave(file.path(fig_rep,"calibration_validation_periods.pdf"), 
       q_hist_plot, width = 10, height = 5)
ggsave(file.path(fig_rep,"calibration_validation_periods.png"), 
       q_hist_plot, width = 10, height = 5)

# Plot on screen
q_hist_plot

# save q_plot dataframe for quick loading in model script
saveRDS(q_plot, file.path(paste0(data_path,"/Discharge"), 
                          "q_cal_val.rds"))
```

# 3 GLACIER DATA

This section processes glacier data for the basin, including volume calculations, spatial analysis, extraction of glacier discharge, and visualization of melt patterns. It handles historical observations and future projections under various climate scenarios, preparing the data for integration with hydrological models.

## 3.1 Background

Note, these code snippets are taken from the script Ex6_Glacier_data.qmd.

For more background, please consult Chapter 6.7 in the CAHAM handbook at https://hydrosolutions.github.io/caham_book/.

We are extracting the glacier runoff from the PyGEM model for the Atbashy basin. The glacier runoff is extracted for each glacier and then aggregated by basin. The glacier runoff is then used as input for the hydrological model. The glacier runoff data can be downloaded from the following publication: Rounce et al. 2023: Global glacier change in the 21st century: Every increase in temperature matters (https://www.science.org/doi/10.1126/science.abo1324)


### Glacier volume calculation using Erasov method

For more information on the Erasov method, please consult the CAHAM handbook at https://hydrosolutions.github.io/caham_book/.


```{r}
# Apply the calculation
results <- calculate_total_glacier_volume(rgi_in_basins)
print(paste("total glacier volume in km^3:", 
    round(sum(results$glacier_volume_km3), 2)))
```


### Remove glacier areas from basin by cropping

Checking if masking was successful by plotting the basin and the masked basin.

```{r}
# remove glacier areas from basin by cropping
basins_masked <- st_difference(basins, st_union(rgi_in_basins))

# validate masking
ggplot() +
  geom_sf(data = basins_masked, fill = "grey", color = "black") +
  theme_minimal() +
  # add title
  labs(title = "Basin Masked by Glacier Outlines") +
  # add north arrow and scale bar
  ggspatial::annotation_north_arrow(location = "tl", 
                                    which_north = "true", pad_x = unit(0.2, "in"), 
                                    pad_y = unit(0.2, "in"), 
                                    style = ggspatial::north_arrow_fancy_orienteering()) +
  ggspatial::annotation_scale(location = "br", width_hint = 0.2)
```

## 3.2 Extract and aggregate glacier discharge for each glacier, aggregate by basin

0.  You need a login for USGS Earth Data here https://ers.cr.usgs.gov/register
1.  Get the download script for the data you require from https://nsidc.org/data/data-access-tool/HMA2_GGP/versions/1
2.  Open a Linux terminal in the folder and run the download script
3.  Open the R project and run the Rmd script to extract mean glacier runoff per basin

Note, in the code below, we have the path to the original Rounce et al. dataset hardwired! This is not good practice but does the job since we have to run the code only once now to get things done.

```{r}
# File to look for
file_name <- "monthly_per_glacier_runoff_river_basin.RData"
full_path <- file.path(glacier_path, file_name)

# Check if the file exists
if (file.exists(full_path)) {
    # Load the RData file
    load(full_path)
    cat("Monthly per glacier runoff data loaded successfully.\n")
    
} else {
    
    # Extracting monthly data from Rounce et al. (2023) dataset
    cat("Monthly per glacier runoff data not found. Extracting data from Rounce et al. (2023) dataset. \n")
    # Data location
    ncfilelist <- list.files(path = "/Users/tobiassiegfried/hydrosolutions Dropbox/Tobias Siegfried/1_HSOL_PROJECTS/PROJECTS/[2020-06-TW] Currricula Strengthening CA/CourseMaterials/Handbook/HydrologicalModeling_CentralAsia_Data/REFERENCE_CaseStudyPacks/CENTRAL_ASIA_DOMAIN/GLACIERS/Rounce_CMIP6_glacier_evolution", pattern = ".*_glac_runoff_fixed_monthly_.*.nc$",  full.names = TRUE)
    
    monthly_per_glacier_runoff <- NULL
    glac_runoff_basin <- NULL
    
    for (ncfile in ncfilelist) {
        
        # Open a file for reading
        file <- ncdf4::nc_open(ncfile)
        # file returns the content of the ncfile. Glacier runoff at a fixed outflow 
        # point is given in m3 each months. 
        
        # Get the RGI IDs and the glacier indices used in the PyGEM data set
        rgiid <- ncdf4::ncvar_get(file, "RGIId")
        
        # Check if the glaciers in the nc files are in one of our basins
        if (sum(rgiid %in% glaciers) == 0) {
            # Glaciers not present in the CA basins
            next
            
        } else {
            
            # Get the rcp identifyer from the file name
            rcp <- str_extract(ncfile, "ssp\\d{3}")
            
            # Get list of models used as climate forcing
            models <- ncdf4::ncvar_get(file, "Climate_Model")
            
            # Get dates for the monthly time series. 
            # The time series starts on January 1, 2000. 
            time <- ncdf4::ncvar_get(file, "time")  # In days
            date <- as_date("2000-01-01") + time
            
            # Get the glacier discharge data
            test <- ncdf4::ncvar_get(file, "glac_runoff_fixed_monthly")
            
            # Calculate mean and std over all climate forcing models
            test_model_mean <- apply(test, c(1,2), mean)
            # test_model_sd <- apply(test, c(1,2), sd)
            
            # Now convert to tibble and long format, filter out the rgiids that are not 
            # located in one of our basins. 
            rownames(test_model_mean) <- paste0(date)
            colnames(test_model_mean) <- paste0(rgiid)
            
            tbl_mean <- as_tibble(test_model_mean, rownames = "date", 
                                  .name_repair = "check_unique") |> 
                dplyr::select(all_of(rgiid[rgiid %in% glaciers])) |> 
                mutate(date = date) |> 
                pivot_longer(-date, names_to = "rgiid", values_to = "qgl_m3month") |> 
                mutate(ssp = rcp)
            
            # Save monthly per-glacier runoff
            monthly_per_glacier_runoff <- rbind(monthly_per_glacier_runoff, tbl_mean)
            
        }
    } 
    
    save(monthly_per_glacier_runoff, file = full_path)
}
```

## 3.3 Compute Basin Aggregates

```{r}
# Check if file exists in directory
if (file.exists(file.path(glacier_path, "monthly_glacier_runoff_agg_basin.RData"))) {
    # Load the RData file
    load(file.path(glacier_path, "monthly_glacier_runoff_agg_basin.RData"))
    cat("Monthly per glacier runoff data loaded successfully.\n")
    
} else {
    cat("Monthly per glacier runoff data not found. Aggregating now. \n")
    monthly_runoff_agg_basin <- monthly_per_glacier_runoff |> 
        left_join(rgi_area_basins |>
                      dplyr::select(RGIId, DnstrNode, weight), 
                  by = c("rgiid" = "RGIId")) 
    
    monthly_runoff_agg_basin2 <- monthly_runoff_agg_basin |> 
        group_by(date, ssp, DnstrNode) |> 
        summarise(glac_runoff_m3month = sum(qgl_m3month * weight)) |> 
        ungroup() |> 
        mutate(model = "mean") |> 
        rename(scenario = ssp) |> 
        drop_na()
    
    # save
    save(monthly_runoff_agg_basin2, 
         file = file.path(glacier_path, 
                          "monthly_glacier_runoff_agg_basin.RData"))
}
```

## 3.4 Norm Glacier Melt

This is used for back extension. We take the norm of the first 5 years of the 21st century, assuming that this is somehow representative for late 20th century glacier melt.

```{r}
#| message: false
#| warning: false

# Calculate average monthly glacier melt between 2000 and 2023
norm_glacier_melt <- monthly_runoff_agg_basin2 |> 
  dplyr::filter(date >= as_date("2000-01-01") & date <= as_date("2004-12-01")) |>
  mutate(month = month(date)) |> 
  group_by(month, DnstrNode) |> 
  summarise(glac_runoff_m3month = mean(glac_runoff_m3month)) |> 
  ungroup() |> 
  mutate(discharge_m3s = glac_runoff_m3month / sec_month)

# Create enhanced plot
pl <- ggplot(norm_glacier_melt, aes(x = month, y = discharge_m3s)) + 
  # Add smoothed line
  geom_line(aes(color = DnstrNode), linewidth = 1) +
  # Add points
  geom_point(aes(color = DnstrNode), size = 2) +
  
  # Customize facets
  facet_wrap(~DnstrNode, scales = "free_y", 
             labeller = labeller(DnstrNode = function(x) paste("Basin", x))) +
  
  # Customize scales
  scale_x_continuous(
    breaks = 1:12,
    labels = month.abb,
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    labels = function(x) format(x, scientific = FALSE, big.mark = ","),
    expand = expansion(mult = c(0.02, 0.1))
  ) +
  scale_color_viridis_d(end = 0.8) +
  
  # Labels
  labs(
    title = "Average Monthly Glacier Melt (2000-2023)",
    subtitle = "Based on monthly runoff aggregation",
    x = "Month",
    y = expression(paste("Discharge [m"^3, "/s]")),
    color = "Basin ID"
  ) +
  
  # Theme customization
  theme_minimal() +
  theme(
    # Text elements
    plot.title = element_text(face = "bold", size = 14, margin = margin(b = 10)),
    plot.subtitle = element_text(color = "grey40", size = 11, margin = margin(b = 20)),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    
    # Panel elements
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey90"),
    panel.spacing = unit(1.5, "lines"),
    
    # Legend
    legend.position = "none",
    
    # Facet labels
    strip.text = element_text(face = "bold", size = 11),
    strip.background = element_rect(fill = "grey95", color = NA),
    
    # Margins
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
  )

# save plot
ggsave(
  filename = file.path(glacier_path, "average_monthly_glacier_melt.png"),
  width = 10, 
  height = 6,
  dpi = 300
)

# Plot on screen
pl
```


## 3.5 Extract Monthly Glacier Discharge from 01/01/2000 through 31/12/2023

We will use airGR for hydrological modeling. Hannah's paper on Atbashy has data analyzed from 2017 - 2024. Let us extract the data for the glacier runoff. We call the period 2000 - 2024 fut_sim.

```{r}
q_glac_2000_2024 <-
   monthly_runoff_agg_basin2 |>
     dplyr::filter(date >= ymd(paste0(hist_obs_start, "-01-01")) & date <= ymd(paste0(fut_sim_end,"-12-31"))) |> dplyr::select(-DnstrNode, -model)

# First calculate the ensemble statistics
q_glac_2000_2024_m3sec <- q_glac_2000_2024 |> 
  group_by(date) |> 
  summarise(
    Glacier = mean(glac_runoff_m3month) / sec_month,
    sd = sd(glac_runoff_m3month) / sec_month
  ) |> 
  ungroup()

# Create enhanced plot
pl <- ggplot(q_glac_2000_2024_m3sec, aes(x = date, y = Glacier)) +
  # Add confidence interval
  geom_ribbon(aes(ymin = Glacier - sd, 
                  ymax = Glacier + sd),
              fill = "steelblue", alpha = 0.2) +
  # Add main line
  geom_line(color = "steelblue", linewidth = 0.8) +
  
  # Customize scales
  scale_x_date(
    date_breaks = "2 years",
    date_labels = "%Y",
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    labels = scales::comma_format(),
    expand = expansion(mult = c(0.02, 0.1))
  ) +
  
  # Labels
  labs(
    title = "Glacier Runoff Time Series (2000-2024)",
    #subtitle = "Ensemble mean with ±1 standard deviation",
    x = "Year",
    y = expression(paste("Discharge [m"^3, "/s]"))#,
    #caption = "Data source: Monthly runoff aggregation"
  ) +
  
  # Theme customization
  theme_minimal() +
  theme(
    # Text elements
    plot.title = element_text(face = "bold", size = 14, margin = margin(b = 10)),
    plot.subtitle = element_text(color = "grey40", size = 11, margin = margin(b = 20)),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1),
    
    # Panel elements
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey90"),
    
    # Plot margins
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
    
    # Caption
    plot.caption = element_text(color = "grey40", margin = margin(t = 20)),
    
    # Background
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

# Save the plot
ggsave(
  filename = file.path(glacier_path, "glacier_runoff_timeseries.png"),
  width = 10, 
  height = 6,
  dpi = 600
)

# Plot on screen
pl

# Save the data
write_csv(q_glac_2000_2024_fut_sim, 
          file = file.path(glacier_path, "q_glac_2000_2024_m3sec.csv"))
```

## 3.6 Back extension of glacier melt to the 1979 to 1996 period

We need glacier melt between 1979 and 1996. We assume the average monthly glacier melt between 2000 and 2004 to have occurred between these years.

```{r}
p_load(numform)

# create date sequence
date_char <- seq(ymd("1979-01-01"), ymd("1999-12-31"), by = "month")
#date_char <- seq(calib_start, valid_end, by = "month")

backtibble <- tibble(
  date = as_date(date_char),
  month = month(date)) |>
  full_join(norm_glacier_melt, by = "month") |>
  dplyr::select(-month) |>
  mutate(glac_runoff_m3s = glac_runoff_m3month/sec_month) |>
  dplyr::select(-glac_runoff_m3month, -discharge_m3s)

glacier_runoff_ssp126 <- backtibble |>
  add_row(
    monthly_runoff_agg_basin2 |>
      dplyr::filter(scenario == "ssp126") |>
      mutate(glac_runoff_m3s = glac_runoff_m3month/sec_month) |>  # s/month
      dplyr::select(-model, -scenario, -glac_runoff_m3month)
    ) |>
  pivot_wider(names_from = DnstrNode, values_from = glac_runoff_m3s,
              names_prefix = "Glacier ")

glacier_runoff_ssp245 <- backtibble |>
  add_row(
    monthly_runoff_agg_basin2 |>
      dplyr::filter(scenario == "ssp245") |>
      mutate(glac_runoff_m3s = glac_runoff_m3month/sec_month) |>  # s/month
      dplyr::select(-model, -scenario, -glac_runoff_m3month)) |>
  pivot_wider(names_from = DnstrNode, values_from = glac_runoff_m3s,
              names_prefix = "Glacier ")

glacier_runoff_ssp370 <- backtibble |>
  add_row(
    monthly_runoff_agg_basin2 |>
      dplyr::filter(scenario == "ssp370") |>      
      mutate(glac_runoff_m3s = glac_runoff_m3month/sec_month) |>  # s/month
      dplyr::select(-model, -scenario, -glac_runoff_m3month)) |>
  pivot_wider(names_from = DnstrNode, values_from = glac_runoff_m3s,
              names_prefix = "Glacier ")

glacier_runoff_ssp585 <- backtibble |>
  add_row(
    monthly_runoff_agg_basin2 |>
      dplyr::filter(scenario == "ssp585") |>
      mutate(glac_runoff_m3s = glac_runoff_m3month/sec_month) |>  # s/month
      dplyr::select(-model, -scenario, -glac_runoff_m3month)) |>
  pivot_wider(names_from = DnstrNode, values_from = glac_runoff_m3s,
              names_prefix = "Glacier ")

glacier_runoff_ssp126 <- glacier_runoff_ssp126 |> 
  mutate(date = floor_date(as.POSIXct(date, tz = "UTC"), unit = "day")) 

glacier_runoff_hist_obs <- glacier_runoff_ssp126 |>  
  dplyr::filter(date >= ymd(paste0(hist_obs_start,"-01-01"))) |>
  dplyr::filter(date <= ymd(paste0(hist_obs_end,"-12-31"))) 

#glacier_runoff_hist_obs <- q_glac_2000_2023_hist_obs

# Assuming glacier_runoff_hist_obs is already defined and loaded with the appropriate data
glaciers_rsm_hist_obs <- process_glacier_runoff(glacier_runoff_hist_obs)

# Save the data
file_name <- "glaciers_hist_obs_rsm.csv"
readr::write_csv(glaciers_rsm_hist_obs, 
                 file.path(glacier_path, file_name), 
                 na = "NA", col_names = FALSE)

# not sure we need this below? Let us check...
# Save to model folder

#glacier_rsm_scen_data_path <- "../05_models/"
#file2write <- file.path(glacier_rsm_scen_data_path, "17084_glaciers_hist_obs_rsm.csv")
#readr::write_csv(glaciers_rsm_hist_obs, pathname, na = "NA", col_names = FALSE)
```

## 3.8 Plot glacier_runoff_hist_obs for cross checking

```{r}
ggplot(glacier_runoff_hist_obs, aes(x = date, y = `Glacier 16076`)) +
  geom_line() +
  labs(title = "Glacier Runoff for historical observation period, 1979 - 1995",
       x = "Date",
       y = "Runoff (m3/s)") +
  theme_minimal() 
```

## 3.9 Future glacier melt

We now save the results for the four climate scenarios: ssp126, ssp245, ssp370, ssp585 Filter from 2000-01-01 to 2024-12-31 and get them in the correct RS MINERVE format.

```{r}
glacier_runoff_fut_ssp126 <- glacier_runoff_ssp126 |>  
  dplyr::filter(date >= ymd(paste0(fut_sim_start,"-01-01"))) |>
  dplyr::filter(date <= ymd(paste0(fut_sim_end,"-12-31"))) |> 
    rename(`Glacier 16076 SSP126` = 2)

# glacier_runoff_ssp245
glacier_runoff_fut_ssp245 <- glacier_runoff_ssp245 |>  
  dplyr::filter(date >= ymd(paste0(fut_sim_start,"-01-01"))) |>
  dplyr::filter(date <= ymd(paste0(fut_sim_end,"-12-31")))|> 
    rename(`Glacier 16076 SSP245` = 2)

# glacier_runoff_ssp370
glacier_runoff_fut_ssp370 <- glacier_runoff_ssp370 |>  
  dplyr::filter(date >= ymd(paste0(fut_sim_start,"-01-01"))) |>
  dplyr::filter(date <= ymd(paste0(fut_sim_end,"-12-31")))|> 
    rename(`Glacier 16076 SSP370` = 2)

glacier_runoff_fut_ssp585 <- glacier_runoff_ssp585 |>  
  dplyr::filter(date >= ymd(paste0(fut_sim_start,"-01-01"))) |>
  dplyr::filter(date <= ymd(paste0(fut_sim_end,"-12-31")))|> 
    rename(`Glacier 16076 SSP585` = 2)

rsm_ssp126 <- process_glacier_runoff(glacier_runoff_fut_ssp126)
rsm_ssp245 <- process_glacier_runoff(glacier_runoff_fut_ssp245)
rsm_ssp370 <- process_glacier_runoff(glacier_runoff_fut_ssp370)
rsm_ssp585 <- process_glacier_runoff(glacier_runoff_fut_ssp585)

readr::write_csv(rsm_ssp126, 
                 file.path(glacier_path,"glaciers_fut_ssp126_rsm.csv"), 
                 na = "NA", col_names = FALSE)
readr::write_csv(rsm_ssp245, 
                 file.path(glacier_path,"glaciers_fut_ssp245_rsm.csv"), 
                 na = "NA", col_names = FALSE)
readr::write_csv(rsm_ssp370, 
                 file.path(glacier_path, "glaciers_fut_ssp370_rsm.csv"), 
                 na = "NA", col_names = FALSE)
readr::write_csv(rsm_ssp585, 
                 file.path(glacier_path,"glaciers_fut_ssp585_rsm.csv"), 
                 na = "NA", col_names = FALSE)
```

## 3.10 Visualization

```{r}
#| message: false
#| warning: false
#| echo: false

# Apply the function to each dataset
hist_long <- prepare_glacier_data_mean_annual_runoff(glacier_runoff_hist_obs |> 
                                                         rename(runoff =2), "obs")
ssp126_long <- prepare_glacier_data_mean_annual_runoff(glacier_runoff_fut_ssp126 |> 
                                                           rename(runoff = 2), "ssp126")
ssp245_long <- prepare_glacier_data_mean_annual_runoff(glacier_runoff_fut_ssp245 |> 
                                                           rename(runoff = 2), "ssp245")
ssp370_long <- prepare_glacier_data_mean_annual_runoff(glacier_runoff_fut_ssp370 |> 
                                                           rename(runoff = 2), "ssp370")
ssp585_long <- prepare_glacier_data_mean_annual_runoff(glacier_runoff_fut_ssp585 |> 
                                                           rename(runoff = 2), "ssp585")

# Combine datasets
combined_data <- bind_rows(hist_long, ssp126_long, ssp245_long, ssp370_long, ssp585_long)

# Create enhanced plot
glacier_runoff_mean_annual <- ggplot(combined_data, 
    aes(x = year, y = runoff, color = scenario, linetype = scenario)) +
    # Add background shading for historical vs future periods
    annotate("rect", xmin = min(combined_data$year), xmax = 2020,
             ymin = -Inf, ymax = Inf, fill = "grey95", alpha = 0.5) +
    
    # Add main lines
    geom_line(linewidth = 1) +
    # add points
    geom_point(size = 1.5, alpha = 0.7) +
    # Add smoothed lines
    geom_smooth(linewidth = .5, method = "loess", span = 0.5, se = FALSE) +
    
    # Customize scales
    scale_x_continuous(
        breaks = seq(min(combined_data$year), max(combined_data$year), by = 10),
        expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(
        labels = scales::comma_format(),
        expand = expansion(mult = c(0.1, 0.1))
    ) +
    # Custom color palette with meaningful colors
    scale_color_manual(
        values = c(
            "obs" = "black",
            "ssp126" = "#2196F3",  # Blue for low emissions
            "ssp245" = "#4CAF50",  # Green for moderate emissions
            "ssp370" = "#FFA000",  # Orange for high emissions
            "ssp585" = "#F44336"   # Red for very high emissions
        ),
        labels = c(
            "obs" = "Historical",
            "ssp126" = "SSP1-2.6 (Low)",
            "ssp245" = "SSP2-4.5 (Moderate)",
            "ssp370" = "SSP3-7.0 (High)",
            "ssp585" = "SSP5-8.5 (Very high)"
        )
    ) +
    # Line types for better distinction
    scale_linetype_manual(
        values = c(
            "obs" = "solid",
            "ssp126" = "solid",
            "ssp245" = "solid",
            "ssp370" = "solid",
            "ssp585" = "solid"
        ),
        labels = c(
            "obs" = "Historical",
            "ssp126" = "SSP1-2.6 (Low)",
            "ssp245" = "SSP2-4.5 (Moderate)",
            "ssp370" = "SSP3-7.0 (High)",
            "ssp585" = "SSP5-8.5 (Very high)"
        )
    ) +
    
    # Labels
    labs(
        title = "Mean Annual Glacier Runoff Projections",
        subtitle = "Historical observations and future scenarios",
        x = "Year",
        y = expression(paste("Mean Annual Discharge [m"^3, "/s]")),
        color = "Scenario",
        linetype = "Scenario"
    ) +
    
    # Theme customization
    theme_minimal() +
    theme(
        # Text elements
        plot.title = element_text(face = "bold", size = 14, margin = margin(b = 10)),
        plot.subtitle = element_text(color = "grey40", size = 11, margin = margin(b = 20)),
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10),
        
        # Legend
        legend.position = "bottom",
        legend.box = "vertical",
        legend.margin = margin(t = 20),
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 10),
        
        # Panel elements
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey90"),
        
        # Plot margins
        plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        
        # Background
        plot.background = element_rect(fill = "white", color = NA),
        panel.background = element_rect(fill = "white", color = NA)
    )

# Save the plot for the report
ggsave(file.path(fig_rep,"glacier_runoff_mean_annual.png"), 
       plot = glacier_runoff_mean_annual, width = 10, height = 5, dpi = 600)

# Plot on screen
glacier_runoff_mean_annual
```

# 4 GENERATION OF HRU

Generates Hydrological Response Units (HRUs) by delineating elevation bands within the basin. The section includes HRU creation, post-processing operations to add attributes like mean elevation and unique identifiers, area calculations, and visualization of the spatial distribution.

## 4.1 HRU Delineation

Note that there are the following options:

-   either work with masked or original basin file
-   either specify the number of elevation bands or the elevation band width (both now work)

For Atbashy, we select the option to work with dh per elevation band.

```{r}
#| message: false
#| warning: false

# configurations
dh_elband <- 200
#n_elband <- 10

# OPTION: select regular basin
basin_utm <- basins |> st_transform(utm42n)
# OPTIONAL select masked basin (i.e., without glaciers)
basin_utm <- basins_masked |> st_transform(utm42n)

# Hypsometric curve based delineation
# mask dem
dem_masked <- dem_utm |> crop(basin_utm) |> mask(basin_utm)
dem_masked_vec <- dem_masked[] |> na.omit()

# ===> OPTION: hardwired elevation range
h_range <- seq(min(dem_masked_vec), max(dem_masked_vec), by = dh_elband) # elevation range steps 
# ===> OPTION: specific number of elevation bands
#h_range <- seq(min(dem_masked_vec), max(dem_masked_vec), length.out = (n_elband + 1)) # elevation range steps 

# classify
dem_utm_class <- classify(dem_utm, h_range)
# polygonize
dem_utm_class_poly <- as.polygons(dem_utm_class)
dem_utm_class_poly_sf <- st_as_sf(dem_utm_class_poly)
dem_utm_class_poly_sf <- dem_utm_class_poly_sf |> 
    mutate(layer = seq(nrow(dem_utm_class_poly_sf)))
# clip and rename
hru_shp_clipped_utm <- 
    st_intersection(dem_utm_class_poly_sf, basin_utm) |> 
    dplyr::select(layer, geometry)
hru <- hru_shp_clipped_utm 

# validation
ggplot() +
  geom_sf(data = hru, fill = "grey", color = "black") +
  theme_minimal() +
  # add title
  labs(title = "Elevation bands") +
  # add north arrow and scale bar
  ggspatial::annotation_north_arrow(location = "tl", which_north = "true", 
    pad_x = unit(0.2, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_fancy_orienteering()) +
  ggspatial::annotation_scale(location = "br", width_hint = 0.2)

# compute number of features (rows) in hru
nrow_hru <- nrow(hru)
print(paste0("Number of Elevation Bands: ", nrow_hru))
```

## 4.2 HRU Post-Processing

### Intersection with Subbasins

This code needs to be used in case we have a more comple basin with n\>1 gauges and the number of associated subbasins. In other words, this code block intersects the elevation bands with subbasins.

!! IMPORTANT -\> RSMINERVE Compliance: Make sure that the basin shapefile contains the subbasin name field as well as the junct_down field as required by RSMINERVE.

#### CASE 1: Single Subbasin

```{r}
hru$name <- river_name
```

#### CASE 2: Multiple Subbasins

Note: this needs to be adjusted when we have multiple subbasins! Check Taloqan code, developed for Wakil Wakil.

### Compute Mean Elevation of HRUs

We add the mean elevation to the HRU. This is required information for RSMinerve.

```{r add_Z_to_hru}
p_load(exactextractr)

zonalStat_Z <- exactextractr::exact_extract(dem_utm,hru,'mean')
hru$Z <- zonalStat_Z
#hru %>% plot()

# Add reversed zonalStat_Z to Basin_Info list and store again on disk.
Basin_Info$ZLayers <- zonalStat_Z |> rev()
saveRDS(Basin_Info, file.path(data_path, "Basin_Info.rds"))
```

### Add unique HRU Names

We divide the river basin to subcatchments and HRUs there. These HRUs are ranked according to mean elevation and numbered accordingly.

```{r add_unique_hru_names}
# unique names in hru
subbasin_names <- unique(hru$name)

# loop through subbasin_names and add unique numbers to corresponding names
for (idx in length(subbasin_names)) {
  subbasin_sel <- hru |> dplyr::filter(name == subbasin_names[idx]) |> dplyr::arrange(Z)
  subbasin_sel$hru_num <- (1:base::nrow(subbasin_sel))
  subbasin_sel$name <- paste0(subbasin_sel$name,'_',subbasin_sel$hru_num)
  
  if (idx == 1) {
    res_subbasins <- subbasin_sel
  } else {
    res_subbasins <- res_subbasins %>% dplyr::add_row(subbasin_sel)
  }
}

res_subbasins$code = gauge_code

hru_named_utm <- res_subbasins %>% dplyr::select(-layer,-hru_num)
#hru_named_utm %>% plot()
hru_named_latlon <- hru_named_utm |> st_transform(crs = crs_project) |> st_make_valid()

# convert name to factor
hru_named_utm$name <- factor(hru_named_utm$name, levels = paste0(paste0(river_name,"_"), 1:nrow_hru))
hru_named_latlon$name <- factor(hru_named_latlon$name, levels = paste0(paste0(river_name,"_"), 1:nrow_hru))
```

### Add HRU areas and X,Y coordinates

```{r}
hru_utm <- post_process_hru(hru_named_utm)
hru_latlon <- post_process_hru(hru_named_latlon)

# preview hru_utm
hru_utm |> 
  dplyr::select(name, area, Z, X, Y) |> 
  dplyr::arrange(name) |> 
  head(10)
```

Improved plotting.

```{r}
# Create the plot with custom colors
pl_elevation_bands <- hru_utm |>
  ggplot() +
  # HRU polygons with improved aesthetics
  geom_sf(aes(fill = name),
          color = "white",
          linewidth = 0.05) +
  # Glacier outlines with improved visibility
  geom_sf(data = rgi_in_basins,
          fill = "transparent",
          color = "lightgrey",
          linewidth = 0.01,
          linetype = "solid") +
  # Add basin outline
  geom_sf(data = basin_latlon,
          fill = "transparent",
          color = "darkgrey",
          linewidth = 0.3) +
  # Use custom color scale
  scale_fill_manual(
    values = atbashy_colors,
    name = "HRU Name"
  ) +
  # Improved theme with better typography
  theme_void() +
  theme(
    # Panel customization
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    # Text improvements
    text = element_text(family = "Arial", color = "grey20"),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    # Legend customization
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 9),
    # Axis labels
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9)
  ) +
  # Add informative labels
  labs(
    title = "Hydrological Response Units with Glacier Outlines",
    subtitle = "Spatial distribution of HRUs and glacier boundaries",
    #caption = "Data source: HRU shapefile and RGI glacier inventory"
  ) +
  # Add scale bar and north arrow
  ggspatial::annotation_scale(
    location = "br",
    width_hint = 0.2
  ) +
  ggspatial::annotation_north_arrow(location = "tl", 
                                    which_north = "true", pad_x = unit(0.2, "in"), 
                                    pad_y = unit(0.2, "in"), 
                                    style = ggspatial::north_arrow_fancy_orienteering())

# store elevation band plot
ggsave(file.path(fig_rep, "elevation_bands.png"),
       plot = pl_elevation_bands,
       width = 10,
       height = 5,
       dpi = 600)

# Plot on screen
pl_elevation_bands <- pl_elevation_bands + 
  theme(text = element_text(family = "sans"))
pl_elevation_bands
```

## 4.3 Save Results

```{r store_shapefile_results}
sf::st_write(hru_utm,file.path(gis_path,"hru_utm.shp"), append = FALSE)
sf::st_write(hru_latlon,file.path(gis_path,"hru_latlon.shp"), append = FALSE)
```

# 5 PROCESSING CLIMATE DATA

This extensive section processes climate data from various sources for both historical observations and future projections. It handles temperature and precipitation data, creates formatted files for modeling software, validates precipitation data against observed lapse rates, and performs necessary transformations for hydrological modeling.

## 5.1 HRU Preparation

See previous section.

## 5.2 Load CHELSA V21 data from the online repository

Data are stored in the repository from 1979 until 1999. This covers the historical observation period which is set from 1979 - 1996.

```{r}
#| message: false
#| warning: false

# if hist_obs_dir is empty, execute the code block below, else print "CHELSA V21 climate data available"
if (length(list.files(hist_obs_dir)) == 0) {
    # create directory if it does not exist
    dir.create(hist_obs_dir, showWarnings = FALSE)
    # download data
    # see also: https://gitlabext.wsl.ch/karger/rchelsa
    extent <- sf::st_bbox(basin_latlon)
    extent <- c(extent$xmin, extent$xmax, extent$ymin, extent$ymax) |> unname()
    
    startdate <- ymd(paste(hist_obs_start,"-1-1"))
    enddate <- startdate + years(1) - days(1)
    # number of years to fetch
    n_years <- 24
    
    for (idx in seq(n_years)) {
        year_name <- year(startdate) |> as.character()
        # temperature tas (note: this is in Kelvin, we need to convert to Celsius.)
        tas <- Rchelsa::getChelsa('tas',extent = extent, startdate = startdate, enddate = enddate)
        tas <- tas - 273.15
        # precipitation pr
        pr <- Rchelsa::getChelsa('pr',extent = extent, startdate = startdate, enddate = enddate)
        # if not already existing, save tas and pr data on disk
        if (!file.exists(file.path(hist_obs_dir,paste0("tas_",year_name,".nc")))) {
            writeCDF(tas, file.path(hist_obs_dir,paste0("tas_",year_name,".nc")),overwrite=TRUE)}
        if (!file.exists(file.path(hist_obs_dir,paste0("pr_",year_name,".nc")))) {
            writeCDF(pr, file.path(hist_obs_dir,paste0("pr_",year_name,".nc")),overwrite = TRUE)}
        # shift dates
        startdate <- enddate + days(1)
        enddate <- startdate + years(1) - days(1)
    }
} else {
    print("CHELSA V21 climate data available - Check years!")
}
```

This is just to validate data download. It all looks good so far.

```{r}
year_name <- 1979

# validate data
tas_2019 <- terra::rast(file.path(hist_obs_dir,paste0("CHELSA_tas__",
                                                      year_name,"_55_85_30_50_V.2.1.nc")))
# extract first day
tas_2019_1 <- tas_2019[[1]]

p_load(tidyterra) # what a glorious package, no more converting to dataframe before plotting with ggplot2.
# plot
ggplot() +
  geom_spatraster(data = tas_2019_1) +
  scale_fill_viridis_c(name = "Temp (°C)") +
  coord_sf()
```

## 5.3 hist_obs: Temperature Data Processing

```{r hist_obs_Climate_Data_Processing}
#| message: false
#| warning: false
#| echo: false

# Parameters
climate_data_type <- "hist_obs"

# Load HRU shapefile (see section 4.0 above)
hru_utm <- sf::st_read(file.path(gis_path,"hru_utm.shp"))

# list all climate files
climate_files_tas <- list.files(hist_obs_dir, pattern = "tas_", full.names = TRUE)
# only process files up to and including hist_obs_end
filtered_climate_files_tas <- climate_files_tas |> 
    keep(~ extract_year(.x) <= hist_obs_end) |> 
    keep(~ extract_year(.x) >= hist_obs_start)

# Data type
temp_or_precip <- "Temperature"

# Do the magic
hist_obs_tas <- 
    gen_HRU_Climate_CSV_RSMinerve_local(filtered_climate_files_tas,
        river_name, temp_or_precip, hru_utm, hist_obs_start,
        hist_obs_end, obs_freq, climate_data_type, crs_project)
```


## 5.4 hist_obs: Precipitation Data Processing

```{r}
#| message: false
#| warning: false
#| echo: false

# list all climate files
climate_files_pr <- list.files(hist_obs_dir, pattern = "pr_", full.names = TRUE)

# only process files up to and including hist_obs_end
filtered_climate_files_pr <- climate_files_pr |> 
    keep(~ extract_year(.x) <= hist_obs_end) |> 
    keep(~ extract_year(.x) >= hist_obs_start)

# Data type
temp_or_precip <- "Precipitation"

# Do the magic
hist_obs_pr <- gen_HRU_Climate_CSV_RSMinerve_local(filtered_climate_files_pr,
        river_name, temp_or_precip, hru_utm, hist_obs_start,
        hist_obs_end, obs_freq, climate_data_type, crs_project)
```

Given the our experience with the very problematic CHELSA V21 forcing in the Zarafshan River basin, we introduce a new Section here for the validation of the precipitation data with observed lapse rate data. This is done by comparing the precipitation data of each elevation band with observed precipitation lapse rates that are obtained from the Soviet compendium on Water Resources of the Soviet Union Amu Darya and Syr Darya (Andrey Yakovlev, Surface water resources. Volume 14 Central Asia, Issue 3 Amudariya).

Note: For Atbashy not implemented.

Below, we compute the elevation band specific norm precipitation (in mm) and the basin average norm precipitation (in mm). The latter is computed by multiplying the elevation band specific norm precipitation with the corresponding elevation band area and dividing it by the total area of the basin without glaciers.

```{r}
#| message: false
#| warning: false
#| echo: false

eb_area <- hru_utm |> dplyr::select(name, area) |> sf::st_drop_geometry() |> type_convert()
area_no_glaciers <- hru_utm$area |> as.numeric() |> sum() # this is in m^2

data_P <- hist_obs_pr |>
    slice(-c(1:8)) |>
    rename(date = Station) |>
    mutate(date = str_sub(date, start = 1, end = 11)) |>
    mutate(date = dmy(date)) |> 
    type_convert()

data_P_long <- data_P |>
    pivot_longer(cols = -date, names_to = "station", values_to = "value") |>
    mutate(sensor = "P") 

# Compute total norm precipitation for each elevation band separately (mm / yr)
pr_norm_eb <- data_P_long |> 
    filter(sensor == "P") |> 
    group_by(station) |> 
    timetk::summarize_by_time(.date_var = date, .by = "1 year", value = sum(value)) |> 
    summarize(value = mean(value))

# now compute the basin average by multiplying each elevation band value by the corresponding elevation band area in eb_area and then summing the result and dividing it by the total area.

pr_norm_basin <- pr_norm_eb |> 
    left_join(eb_area, by = c("station" = "name")) |> 
    mutate(value = value * area) |> 
    summarize(value = sum(value) / area_no_glaciers) |> 
    rename(pr_norm_basin = value)
pr_norm_basin

pr_norm_eb
```


Let us plot the elevation gradient of the precipitation for each elevation band. The data is in pr_norm_eb. 

```{r}
pl_pr_norm_eb <- pr_norm_eb |> 
    mutate(
        # Convert station to factor with correct order
        station = factor(station, 
                        levels = paste0("Atbashy_", 1:11)),
        # Add labels for each bar
        value_label = sprintf("%.0f", value)
    ) |>
    ggplot(aes(x = station, y = value, fill = station)) +
    # Add columns with custom colors
    geom_col(width = 0.7) +
    # Add value labels on top of each bar
    geom_text(aes(label = value_label),
              position = position_dodge(width = 0.7),
              vjust = -0.5,
              size = 3) +
    # Add basin average line
    geom_hline(yintercept = pr_norm_basin$pr_norm_basin, 
               linetype = "dashed", 
               color = "navy",
               linewidth = 0.7) +
    # Add basin average label
    annotate("text", 
             x = 1, 
             y = pr_norm_basin$pr_norm_basin,
             label = sprintf("Basin Average: %.0f mm", pr_norm_basin$pr_norm_basin),
             hjust = -0.1,
             vjust = -0.5,
             color = "navy",
             fontface = "bold") +
    # Use custom color scale
    scale_fill_manual(values = atbashy_colors) +
    # Customize labels
    labs(title = paste(river_name, "Precipitation Gradient"),
         subtitle = "Based on CHELSA V2.1 Data",
         x = "Elevation Band",
         y = "Annual Precipitation (mm)") +
    # Improve theme
    theme_minimal() +
    theme(
        # Text improvements
        text = element_text(family = "Arial"),
        plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        axis.text.y = element_text(size = 9),
        # Remove legend
        legend.position = "none",
        # Grid customization
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(color = "grey90")
    ) +
    # Expand y-axis slightly to accommodate labels
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)))

# save plot
ggsave(file.path(fig_rep, "precipitation_elevation_gradient.png"),
       plot = pl_pr_norm_eb,
       width = 10,
       height = 5,
       dpi = 600)

# plot on screen
pl_pr_norm_eb
```

Do the same polished plot with Z on the x-axis, i.e. the mean elevation band elevation.

```{r}
#| echo: false
#| message: false
#| warning: false

# prepare elevation band areas (in m2) 
eb_area <- hru_utm |> dplyr::select(name, area) |> sf::st_drop_geometry() |> type_convert()
area_no_glaciers <- hru_utm$area |> as.numeric() |> sum() # this is in m^2

#eb_Z <- hru_utm |> dplyr::select(name, Z) |> sf::st_drop_geometry() |> type_convert()
#eb_Z$Z <- eb_Z$Z |> round()

# add Z from eb_Z to pr_norm_eb
eb_Z <- hru_utm |> dplyr::select(name, Z) |> sf::st_drop_geometry() |> type_convert()
eb_Z$Z <- eb_Z$Z |> round()

pr_norm_Z_eb <- pr_norm_eb |> 
    left_join(eb_Z, by = c("station" = "name")) |>
    arrange(Z)

# Create an improved precipitation-elevation plot
pl_pr_norm_Z_eb <- pr_norm_Z_eb |> 
    # Add station labels for the points
    mutate(
        station_label = str_replace(station, "Atbashy_", ""),
        value_label = sprintf("%.0f mm", value)
    ) |>
    ggplot(aes(x = Z, y = value)) +
    # Add gridlines behind the data
    theme_minimal() +
    # Add main line with custom styling
    geom_line(linewidth = 0.8, 
              color = "navy") +
    # Add points with custom styling
    geom_point(aes(fill = station),
               size = 3,
               shape = 21,
               color = "white",
               stroke = 0.5) +
    # Add basin average reference line
    geom_hline(yintercept = pr_norm_basin$pr_norm_basin, 
               linetype = "dashed", 
               color = "navy",
               linewidth = 0.7,
               alpha = 0.5) +
    # Add basin average label
    annotate("text", 
             x = min(pr_norm_Z_eb$Z), 
             y = pr_norm_basin$pr_norm_basin,
             label = sprintf("Basin Average: %.0f mm", pr_norm_basin$pr_norm_basin),
             hjust = 0,
             vjust = -0.5,
             color = "navy",
             fontface = "bold") +
    # Add point labels
    geom_text(aes(label = station_label),
              size = 3,
              vjust = -1.5) +
    # Use custom color scale matching previous plots
    scale_fill_manual(values = atbashy_colors) +
    # Customize axis scales
    scale_x_continuous(
        breaks = seq(floor(min(pr_norm_Z_eb$Z)/500)*500, 
                    ceiling(max(pr_norm_Z_eb$Z)/500)*500, 
                    by = 500),
        labels = function(x) sprintf("%d m", x)
    ) +
    # Increased frequency of y-axis breaks (changed from 500 to 250)
    scale_y_continuous(
        breaks = seq(0, ceiling(max(pr_norm_Z_eb$value)/250)*250, by = 250),
        labels = function(x) sprintf("%d mm", x),
        expand = expansion(mult = c(0.02, 0.1))
    ) +
    # Add labels
    labs(title = "Precipitation-Elevation Relationship",
         subtitle = paste(river_name, "- Based on CHELSA V2.1 Data"),
         x = "Elevation (masl)",
         y = "Annual Precipitation (mm/a)") +
    # Customize theme
    theme(
        text = element_text(family = "Arial"),
        plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10, face = "bold"),
        axis.text = element_text(size = 9),
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey90")
    )

# save figure
ggsave(file.path(fig_rep, "precipitation_elevation_relationship.png"),
       plot = pl_pr_norm_Z_eb,
       width = 10,
       height = 5,
       dpi = 600)

# Plot on screen
pl_pr_norm_Z_eb
```


### hist_obs: Precipitation Rescaling? Y/N

Here, we add the option to rescale precipitation data. This is done by multiplying the precipitation data of each elevation band by a factor. This is useful when we want to adjust the precipitation reanalysis data of each elevation band to better match observed precipitation lapse rates.

```{r}
# rescale precipitation data? Y/N
rescale_pr <- "N"

if (rescale_pr == "Y") {
    # extract the header from hist_obs_pr (first seven rows)
    hist_obs_pr_header <- hist_obs_pr[1:7,]
    # extract the data from hist_obs_pr (from row 8 onwards) and convert to dbl.
    hist_obs_pr_data <- hist_obs_pr[8:nrow(hist_obs_pr),] |> type_convert()
    # extract Station column
    hist_obs_pr_station <- hist_obs_pr_data |> dplyr::select(Station)
    hist_obs_pr_data <- hist_obs_pr_data |> dplyr::select(-Station)
    # multiply each values in each column with the matching corresponding scaling factor from pr_norm_Z_eb
    hist_obs_pr_rescaled <- hist_obs_pr_data / pr_norm_Z_eb$scaling_factor
    hist_obs_pr_data
    hist_obs_pr_rescaled
    
    #reassemble all into hist_obs_pr_rescaled
    hist_obs_pr_rescaled <- hist_obs_pr_station |> 
        add_column(hist_obs_pr_rescaled, .name_repair = "unique")
    # add header on top
    hist_obs_pr_rescaled <- rbind(hist_obs_pr_header, hist_obs_pr_rescaled)
} else {
    hist_obs_pr_rescaled <- hist_obs_pr
}
```

## 5.5 hist_obs: Combine All Data and Add Discharge

```{r}
#| warning: false

forcing_dir <- paste(model_dir, "forcing", sep = "/")

# Combine extracted climate tibbles
hist_obs_rsm <- hist_obs_tas |> 
  add_column(hist_obs_pr %>% dplyr::select(-Station),.name_repair = 'unique')

if (rescale_pr == "Y") {
    hist_obs_rsm_rescaled <- hist_obs_tas |> 
        add_column(hist_obs_pr_rescaled %>% dplyr::select(-Station),.name_repair = 'unique')
}

combined_header <- c(names(hist_obs_tas), rbind(names(hist_obs_pr) %>% .[-1]))

if (hydro_model_time_steps == "mon") {q_obs_sim <- q_mon
} else if (hydro_model_time_steps == "dec") {q_obs_sim <- q_dec 
} else if (hydro_model_time_steps == "day") {q_obs_sim <- q_dly }

q_obs_sim <- q_obs_sim |> mutate(name = "value") |> 
    dplyr::select(date,name,value) |> 
    dplyr::filter(date >= ymd(paste0(hist_obs_start,"-01-01"))) |> 
    dplyr::filter(date <= ymd(paste0(hist_obs_end,"-12-31"))) |> 
    mutate(value = as.character(value))

q_obs_sim_wide <- q_obs_sim |> 
    pivot_wider(values_from = value, names_from = name) |> 
    arrange(date) |> 
    rename(discharge = value)

dates_char_Q <- 
    posixct2rsminerveChar(q_obs_sim_wide$date, tz = "UTC") |>  
    rename(Station = value) |> 
    tibble::add_column(q_obs_sim_wide |> dplyr::select(-date))

combined_header <- c(combined_header, names(q_obs_sim_wide) %>% .[-1])

# Combine everything
hist_obs_rsm <- dplyr::full_join(hist_obs_rsm, dates_char_Q, by = 'Station')

if (rescale_pr == "Y") {
    hist_obs_rsm_rescaled <- dplyr::full_join(hist_obs_rsm_rescaled,
                                              dates_char_Q_mon, by = 'Station') |> 
        dplyr::full_join(dates_char_Q_dec, by = 'Station')
}

# now finish off by giving the required attributes in the table for the discharge station
## monthly data
hist_obs_rsm$discharge[1] = gauge_coord[1]
hist_obs_rsm$discharge[2] = gauge_coord[2]
hist_obs_rsm$discharge[3] = gauge_Z 
hist_obs_rsm$discharge[4] = "Q"
hist_obs_rsm$discharge[5] = 'Flow'
hist_obs_rsm$discharge[6] = "m3/s"
hist_obs_rsm$discharge[7] = 'Constant before'

# do the same for the rescaled dataframe
if (rescale_pr == "Y") {
    hist_obs_rsm_rescaled$discharge[1] = gauge_coord[1]
    hist_obs_rsm_rescaled$discharge[2] = gauge_coord[2]
    hist_obs_rsm_rescaled$discharge[3] = gauge_Z 
    hist_obs_rsm_rescaled$discharge[4] = "Q"
    hist_obs_rsm_rescaled$discharge[5] = 'Flow'
    hist_obs_rsm_rescaled$discharge[6] = "m3/s"
    hist_obs_rsm_rescaled$discharge[7] = 'Constant before'
}

# write to disk
file2write <- base::rbind(combined_header,hist_obs_rsm)
readr::write_csv(file2write, 
                 file.path(model_dir,"forcing/hist_obs_rsm.csv"), 
                 na = "NA", col_names = FALSE)

# rescaled version
if (rescale_pr == "Y") {
    file2write <- base::rbind(combined_header,hist_obs_rsm_rescaled)
    readr::write_csv(file2write, file.path(forcing_dir,"hist_obs_rsm_rescaled.csv"), na = "NA", col_names = FALSE)
}
```

## 5.6 hist_sim: Historical ERA5 Simulations over hist_obs period (1979 - 1996)

hist_sim is ERA5 data from 1979 - 1996.

```{r}
#| warning: false
#| echo: false
#| message: false

# Parameters
climate_data_type <- "hist_sim"

# read hist_sim file from ./02_data/Forcing/ERA5/ to forcing_dir and rename to hist_sim_ERA5_1979_1996.csv
fut_sim_RSM <- read_csv('../02_data/Forcing/ERA5/HRU_16076Atbashy_1979-01-01_1996-12-31.csv')

# Write result to disk
file2write <- base::rbind(head(combined_header,-1),fut_sim_RSM)
write_csv(file2write, paste0(forcing_dir,"/",climate_data_type,"_",hist_sim_models,"_",
          gauge_code,"_",hist_sim_start,"_",hist_sim_end,".csv"), col_names = FALSE)
```


<!-- ```{r hist_sim_Climate_Data_Processing} -->
<!-- #| warning: false -->
<!-- #| echo: false -->
<!-- #| message: false -->

<!-- # Parameters -->
<!-- climate_data_type <- "hist_sim" -->
<!-- climate_files_tas <- list.files(hist_sim_dir,pattern = "tas_", full.names = TRUE) -->
<!-- climate_files_pr <- list.files(hist_sim_dir,pattern = "pr_", full.names = TRUE) -->

<!-- # Extract GCM Model-Specific Data -->
<!-- hist_sim_rsm <- vector(mode = "list", length = length(gcm_Models)) -->
<!-- names(hist_sim_rsm) <- gcm_Models -->

<!-- for (idxGCM in seq(length(gcm_Models))) { -->

<!--   temp_or_precip <- "Temperature" -->
<!--   gcm_model <- gcm_Models[idxGCM] -->
<!--   hist_sim_T <- gen_HRU_Climate_CSV_RSMinerve_local(climate_files_tas[idxGCM], -->
<!--                                               river_name, -->
<!--                                               temp_or_precip, -->
<!--                                               hru_utm, -->
<!--                                               hist_sim_start, -->
<!--                                               hist_sim_end, -->
<!--                                               obs_freq, -->
<!--                                               climate_data_type, -->
<!--                                               crs_project) -->

<!--   temp_or_precip <- "Precipitation" -->
<!--   hist_sim_P <- gen_HRU_Climate_CSV_RSMinerve_local(climate_files_pr[idxGCM], -->
<!--                                               river_name, -->
<!--                                               temp_or_precip, -->
<!--                                               hru_utm, -->
<!--                                               hist_sim_start, -->
<!--                                               hist_sim_end, -->
<!--                                               obs_freq, -->
<!--                                               climate_data_type, -->
<!--                                               crs_project) -->

<!--   hist_sim_rsm[[idxGCM]] <- hist_sim_T %>%  -->
<!--     tibble::add_column(hist_sim_P %>% dplyr::select(-Station),.name_repair = 'unique') -->

<!--   # Write result to disk -->
<!--   file2write <- base::rbind(combined_header,hist_sim_rsm[[idxGCM]]) -->
<!--   write_csv(file2write,paste0(forcing_dir,"/",climate_data_type,"_",gcm_model,"_", -->
<!--             gauge_code,"_",hist_sim_start,"_",hist_sim_end,".csv"), col_names = FALSE) -->
<!-- } -->
<!-- ``` -->


## 5.7 fut_sim: 'Future' ERA5 Simulations 2000 - 2023

The forcing data has been obtained from the data gateway from 2000 - end of 2024.
For proper bias correction, we should actually get the data from 1979 onward so that we can do the 
bias correction properly between CHELSA V21 and ERA5.

Below, we just read the output from the data gateway and write it to disk.

```{r}
#| warning: false
#| echo: false
#| message: false

# dealing with 'future simulations'
climate_data_type <- "fut_sim"

# read file from ./02_data/Forcing/ERA5/ to forcing_dir and rename to fut_sim_ERA5_2000_2024.csv
fut_sim_RSM <- read_csv('../02_data/Forcing/ERA5/HRU_16076Atbashy_2000-01-01_2024-12-31.csv')

# Write result to disk
file2write <- base::rbind(head(combined_header,-1),fut_sim_RSM)
write_csv(file2write, paste0(forcing_dir,"/",climate_data_type,"_",hydrology_run_Models_Scenarios,"_",
          gauge_code,"_",fut_sim_start,"_",fut_sim_end,".csv"), col_names = FALSE)
```


# the code below is not needed as we do not carry out GCM simulations
<!-- ```{r fut_sim_Climate_Data_Processing, echo=FALSE, warning=FALSE} -->
<!-- #| warning: false -->
<!-- #| echo: false -->
<!-- #| message: false -->

<!-- climate_data_type <- "fut_sim" -->

<!-- # Process and Extract GCM Model-Specific Data -->
<!-- fut_sim_rsm <- base::vector(mode = "list", length = length(gcm_Models_Scenarios)) -->
<!-- names(fut_sim_rsm) <- gcm_Models_Scenarios # now we have a named list -->

<!-- debug_T <- fut_sim_rsm -->

<!-- for (idxGCM in seq(length(gcm_Models_Scenarios))) { -->

<!--   # GCM Model and Scenario -->
<!--   str2proc <- gcm_Models_Scenarios[idxGCM] -->
<!--   gcm_model <- substr(str2proc, 1, nchar(str2proc) - 7) -->
<!--   gcm_scenario <- substr(str2proc, nchar(str2proc) - 6 + 1, nchar(str2proc)) -->

<!--   # Process files - tas -->
<!--   temp_or_precip <- "Temperature" -->
<!--   climate_file_tas <-  -->
<!--     list.files(fut_sim_dir, -->
<!--                pattern = paste0("tas_day_",gcm_Models_Scenarios[idxGCM]), -->
<!--                full.names = TRUE) -->
<!--   fut_sim_T <- gen_HRU_Climate_CSV_RSMinerve_local(climate_file_tas, -->
<!--                                              river_name, -->
<!--                                              temp_or_precip, -->
<!--                                              hru_utm, -->
<!--                                              fut_sim_start, -->
<!--                                              fut_sim_end, -->
<!--                                              obs_freq, -->
<!--                                              climate_data_type, -->
<!--                                              crs_project) -->

<!--   debug_T[[idxGCM]] <- fut_sim_T -->

<!--   # Process files - pr -->
<!--   temp_or_precip <- "Precipitation" -->
<!--   climate_file_pr <-  -->
<!--     list.files(fut_sim_dir, -->
<!--                pattern = paste0("pr_day_",gcm_Models_Scenarios[idxGCM]), -->
<!--                full.names = TRUE) -->
<!--   fut_sim_P <- gen_HRU_Climate_CSV_RSMinerve_local(climate_file_pr, -->
<!--                                              river_name, -->
<!--                                              temp_or_precip, -->
<!--                                              hru_utm, -->
<!--                                              fut_sim_start, -->
<!--                                              fut_sim_end, -->
<!--                                              obs_freq, -->
<!--                                              climate_data_type, -->
<!--                                              crs_project) -->

<!--   # Final dataframe -->
<!--   fut_sim_rsm[[idxGCM]] <- fut_sim_T %>%  -->
<!--     tibble::add_column(fut_sim_P %>% dplyr::select(-Station),.name_repair = 'unique') -->
<!--   # Write result to disk -->
<!--   file2write <- base::rbind(combined_header,fut_sim_rsm[[idxGCM]]) -->
<!--   readr::write_csv(file2write,paste0(forcing_dir,'/',climate_data_type,"_",gcm_model,"_", -->
<!--                    gcm_scenario,"_",fut_sim_start,"_",fut_sim_end,".csv"), -->
<!--                    col_names = FALSE) -->
<!-- } -->
<!-- ``` -->


# 6 BIAS CORRECTION OF fut_sim

Applies quantile mapping bias correction to future climate simulations, ensuring consistency between historical observations and future projections. This process is critical for reliable climate change impact assessment.

## 6.1 Actual Bias Correction

```{r fut_sim_bcsd, message = FALSE, warning = FALSE}
# Quantile mapping package
p_load(qmap)

# Preparations
## HRUs
n_hru <- hru_utm |> nrow()
hru_names <- hru_utm$name

# ================
# Prepare hist_obs
# ================
climate_data_type <- "hist_obs"
#####################################################################
#### CHANGE HERE IF ORIGINAL OR RESCALED HIST_OBS DATA SHOULD BE USED
if (rescale_pr == "Y") {
  hist_obs_path <- file.path(forcing_dir,"hist_obs_rsm_rescaled.csv")
} else {
  hist_obs_path <- file.path(forcing_dir,"hist_obs_rsm.csv")
}
#####################################################################
hist_obs_orig <- hist_obs_path |>  readr::read_csv(col_types = cols(.default = col_character())) 
# delete first and last column of hist_obs_orig (or. more general, delete first column and all the discharge observations)
hist_obs_orig <- hist_obs_orig |> dplyr::select(-1,ncol(hist_obs_orig))

# Extract data by groups and convert T to deg. K
hist_obs_T <- hist_obs_orig[,1:n_hru] |> 
    slice(-1:-7) |> 
    type_convert() |> 
    mutate(across(.cols = everything(), ~ . + 273.15))
hist_obs_P <- hist_obs_orig[,(n_hru + 1):(2 * n_hru)] |> slice(-1:-7) |> type_convert()

# Fix row names
names(hist_obs_T) <- hru_names
names(hist_obs_P) <- hru_names

hist_obs_T_df <- hist_obs_T %>% as.data.frame()
row.names(hist_obs_T_df) <- hist_obs_dates$Date %>% as.character()
hist_obs_P_df <- hist_obs_P %>% as.data.frame()
row.names(hist_obs_P_df) <- hist_obs_dates$Date %>% as.character()

# ================
# Prepare hist_sim
# ================
hist_sim_T_list <- base::vector(mode = "list", length = length(hist_sim_models))
hist_sim_P_list <- base::vector(mode = "list", length = length(hist_sim_models))
names(hist_sim_T_list) <- hist_sim_models # now we have a named list
names(hist_sim_P_list) <- hist_sim_models # now we have a named list

for (idx in seq(length(hist_sim_models))) {
  hist_sim_path <- 
    list.files(paste0(model_dir,"/forcing"),
               pattern = paste0("hist_sim_",hist_sim_models[idx]),
               full.names = TRUE)
  hist_sim_orig <- hist_sim_path |> 
      readr::read_csv(col_types = cols(.default = col_character())) |> 
      dplyr::select(-Station)
  
  hist_sim_T <- hist_sim_orig[,1:n_hru] |> 
      slice(-1:-7) |> 
      type_convert() |> 
      mutate(across(.cols = everything(), ~ . + 273.15))
  hist_sim_P <- hist_sim_orig[,(n_hru + 1):(2 * n_hru)] |> 
      slice(-1:-7) |> 
      type.convert()
  
  # Fix row names
  names(hist_sim_T) <- hru_names
  names(hist_sim_P) <- hru_names
  
  hist_sim_T_df <- hist_sim_T %>% as.data.frame()
  row.names(hist_sim_T_df) <- hist_sim_dates$Date %>% as.character()
  hist_sim_P_df <- hist_sim_P %>% as.data.frame()
  row.names(hist_sim_P_df) <- hist_sim_dates$Date %>% as.character()
  
  hist_sim_T_list[[idx]] <- hist_sim_T_df
  hist_sim_P_list[[idx]] <- hist_sim_P_df
}

# ===============
# Prepare fut_sim
# ===============
fut_sim_T_list <- base::vector(mode = "list", length = length(hydrology_run_Models_Scenarios))
fut_sim_P_list <- base::vector(mode = "list", length = length(hydrology_run_Models_Scenarios))
names(fut_sim_T_list) <- hydrology_run_Models_Scenarios # now we have a named list
names(fut_sim_P_list) <- hydrology_run_Models_Scenarios
fut_sim_T_list_bcsd <- fut_sim_T_list
fut_sim_P_list_bcsc <- fut_sim_P_list

for (idx in 1:length(hydrology_run_Models_Scenarios)) {
  fut_sim_orig_path <- 
    list.files(forcing_dir,
               pattern = paste0("fut_sim_",hydrology_run_Models_Scenarios[idx]),
               full.names = TRUE)
  fut_sim_orig <- fut_sim_orig_path %>% 
    readr::read_csv(col_types = cols(.default = col_character())) %>% dplyr::select(-Station)
  
  fut_sim_T <- fut_sim_orig[,1:n_hru] %>% slice(-1:-7) %>% 
    mutate(across(everything(), ~ as.numeric(.) + 273.15))
  fut_sim_P <- fut_sim_orig[, (n_hru + 1) : (2 * n_hru)] %>% 
    slice(-1 : -7) %>% mutate_all(as.numeric)
  
  # Fix row names
  names(fut_sim_T) <- hru_names
  names(fut_sim_P) <- hru_names
  
  fut_sim_T_df <- fut_sim_T %>% as.data.frame()
  row.names(fut_sim_T_df) <- fut_sim_dates$Date %>% as.character()
  fut_sim_P_df <- fut_sim_P %>% as.data.frame()
  row.names(fut_sim_P_df) <- fut_sim_dates$Date %>% as.character()
  
  fut_sim_T_list[[idx]] <- fut_sim_T_df
  fut_sim_P_list[[idx]] <- fut_sim_P_df
}

# ===================
# Do quantile mapping
# ===================

# --- Debugging
fut_sim_rsm_qmapped <- base::vector(mode = "list", length = length(hydrology_run_Models_Scenarios))
names(fut_sim_rsm_qmapped) <- hydrology_run_Models_Scenarios # now we have a named list
# -----

for (idx in seq(length(hydrology_run_Models_Scenarios))) {
  
  # Preparation
  str2proc <- hydrology_run_Models_Scenarios[idx]
  gcm_model <- fut_sim_models
  gcm_scenario <- hydrology_run_scenarios
  
  # Bias correction
  hist_sim_T_df_gcmModel <- hist_sim_T_list[[gcm_model]]
  hist_sim_P_df_gcmModel <- hist_sim_P_list[[gcm_model]]
  
  fut_sim_T_df_gcmModel <- fut_sim_T_list[[idx]]
  fut_sim_P_df_gcmModel <- fut_sim_P_list[[idx]]
  
  qmap_param_T_gcm <- fitQmap(hist_obs_T_df, hist_sim_T_df_gcmModel, method = "QUANT", wet.day = FALSE)
  qmap_param_P_gcm <- fitQmap(hist_obs_P_df, hist_sim_P_df_gcmModel, method = "QUANT")
  
  # T bias correction
  fut_sim_T_df_gcmModel_qmapped <- doQmap(fut_sim_T_df_gcmModel,qmap_param_T_gcm)
  # Fill 0s where present. Occasionally, there are 0s resulting from the quantile mapping these 
  # are ironed out here by filling the missing values using the ones from the preceeding observation. 
  fut_sim_T_df_gcmModel_qmapped[fut_sim_T_df_gcmModel_qmapped == 0] <- NA
  fut_sim_T_df_gcmModel_qmapped <- zoo::na.locf(fut_sim_T_df_gcmModel_qmapped)
  # P bias correction
  fut_sim_P_df_gcmModel_qmapped <- doQmap(fut_sim_P_df_gcmModel,qmap_param_P_gcm)

  # go back to tibble and convert back to deg. C
  fut_sim_T_gcmModel_qmapped <- fut_sim_T_df_gcmModel_qmapped %>% as_tibble() %>% 
    add_column(Date = fut_sim_dates$Date,.before = 1)
  fut_sim_T_gcmModel_qmapped <- fut_sim_T_gcmModel_qmapped %>% mutate(across(-Date, ~ . - 273.15))
  fut_sim_P_gcmModel_qmapped <- fut_sim_P_df_gcmModel_qmapped %>% as_tibble() %>% 
    add_column(Date = fut_sim_dates$Date,.before = 1)  
  
  fut_sim_T_list[[idx]] <- fut_sim_T_gcmModel_qmapped
  fut_sim_P_list[[idx]] <- fut_sim_P_gcmModel_qmapped
  
  # Export to .csv-file
  fut_sim_qmapped <- 
    fut_sim_T_gcmModel_qmapped %>% #dplyr::select(-Date) %>% 
    tibble::add_column(fut_sim_P_gcmModel_qmapped %>% dplyr::select(-Date),.name_repair = "universal") %>% 
    mutate(across(.cols = everything(),~ as.character(.))) %>% dplyr::select(-Date)
  
  # --- Debugging
  fut_sim_rsm_qmapped[[idx]] <- fut_sim_qmapped
  # ---
  
  fut_sim_orig_path <- 
    list.files(forcing_dir,
               pattern = paste0("fut_sim_",hydrology_run_Models_Scenarios[idx]),
               full.names = TRUE)
  fut_sim_orig <- 
    fut_sim_orig_path %>% 
    readr::read_csv(col_types = cols(.default = col_character()),col_names = FALSE) 
  
  fut_sim_orig_header <- fut_sim_orig %>% dplyr::select(-X1) %>% dplyr::slice(1:8,) 
  names(fut_sim_qmapped) <- names(fut_sim_orig_header)
  fut_sim_qmapped <- fut_sim_orig_header  %>% bind_rows(fut_sim_qmapped) %>% 
    add_column(Station = fut_sim_orig$X1,.before = 1)

  # Write result to disk
  climate_data_type <- "fut_sim_bcsd"
  
  readr::write_csv(fut_sim_qmapped,
                   paste0(forcing_dir,'/',climate_data_type,"_",gcm_model,"_",
                          gcm_scenario,"_",fut_sim_start,"_",
                          fut_sim_end,".csv"), col_names = FALSE)
}
```

# 7 VALIDATION OF RESULTS

The final section validates the prepared data through visualization of historical data and future projections. It includes plots of temperature and precipitation for different climate models and scenarios, confirming the successful preparation of model inputs.

## 7.1 Historic Data

First, we load the historic precipitation data.

```{r}
# From this file
hist_obs <- read_csv(file.path(forcing_dir,"hist_obs_rsm.csv"), col_names = FALSE)
#delete the last column in hist_obs (discharge measured at the gauge)
hist_obs <- hist_obs %>% dplyr::select(-ncol(hist_obs)) 

# ... to this file
hist_obs_long <- process_rsm_data(hist_obs, "historical") 
hist_obs_long
```

## 7.2 Future Data

### Load BCSD Data of all ERA5 Model

```{r}
#| echo: false
#| message: false
#| warning: false 

# fut_sim
fut_sim_bcsd <- 
  read_csv(file.path(forcing_dir,
                     "fut_sim_bcsd_ERA5_run_2000_2024.csv"), 
           col_names = FALSE)

fut_sim_bcsd_long <- process_rsm_data(fut_sim_bcsd, "hydrology_run")

all_data_era5_bcsd <- bind_rows(era5_bcsd_long) |> mutate(model = "fut_sim")

# hist_obs_rsm
hist_obs_rsm <- 
  read_csv(file.path(forcing_dir,
                     "hist_obs_rsm.csv"), 
           col_names = FALSE) |> 
  # delete last column
  dplyr::select(-ncol(hist_obs_rsm))

hist_obs_rsm_long <- process_rsm_data(hist_obs_rsm, "historical")
hist_obs_rsm_long <- hist_obs_rsm_long |> mutate(model = "hist_obs")

# combine dataframes
df <- bind_rows(all_data_era5_bcsd, hist_obs_rsm_long)
```

### Plot Data, Elevation Band 1

We just make a sample plot to validate the data preparation steps.

```{r}
#| warning: false
#| echo: false
#| message: false

# plot annual mean temperatures (sensor=="T") of df, grouped by model, for station==Atbashy_1. First, compute annual mean, then plot.
df |> 
    filter(station == "Atbashy_1") |> 
    filter(sensor == "T") |> 
    group_by(model) |> 
    timetk::summarize_by_time(.by = "year",value = mean(value)) |> 
    ungroup() |> 
    ggplot(aes(x = date, y = value, color = model)) +
    geom_line() +
    labs(title = paste(river_name, "Temperature at Atbashy_1"),
         x = "Date",
         y = "Temperature (°C)") +
    theme_minimal() +
    theme(text = element_text(family = "Arial"))



```

Looking all good. Now, we can proceed with the hydrological modeling and the climate impact analysis. The results are investigated in the corresponding script.
